Computing volume via oblique or irregularly shaped rectangular infinitesimals

### 2019-08-20 Cav2D/article-template.tex
initial fixes of article

\section{Cavalieri Integral}
The single Cavalieri integral was originally defined by Ackermann et al. without having to resort to vector calculus \cite{ackermann12,stewart1999}. In this section, however, we will define the Cavalieri integral using vector calculus notation. As the reader will undoubtedly realize, if we define the single Cavalieri integral using vector calculus then it becomes trivial to
extend its definition into $\mathbb{R}^3$.\\

@@

\begin{definition}[Translational curve in $\mathbb{R}^2$]
If a curve $\mathbf{c}(t)$ in $\mathbb{R}^2$ intersects the $x$ axis at the origin and $\mathbf{c}(t) + \left < e,0 \right >$ intersects 
the function $f(x)$ at a single point for all $e\in [a,b]$ then $\mathbf{c}(t)$ is a translational curve 
with respect to $f(x)$ on $[a,b]$. 
\end{definition}

@@

\begin{equation}
\label{eq:cav_1d}
\int \limits_{\mathcal{C}} f(x)~dx = \lim_{n \rightarrow \infty}  \sum_{i=0}^{n-1} f(x_i^2) \Delta x_i^1,
\end{equation}
with $\mathcal{C} = \{\mathbf{c}(t),[a,b]\}$. Equation~\eqref{eq:cav_1d} is difficult to evaluate in its current form. Luckily, the Cavalieri integral 
can be converted into an equivalent Riemann integral using the transformation $h$:
 \begin{equation}
 \int \limits_{\mathcal{C}} f(x)\, dx = \int_{a}^{b} f\circ h(x)~dx. 
\end{equation}

\noindent
The only thing that now remains is to find an expression for $h$. Note that, for every coordinate $(x_i^1,x_i^2) \in [a,b] \times [a',b']$ there exists a unique $t\in\mathbb{R}$ such that
\begin{equation}
 \label{eq:pat_rel_1d}
 \left <x_i^2,f(x_i^2) \right > = \left <x_i^1,0 \right> + \left<c_x(t),c_y(t) \right >.
 \end{equation}
 We can rewrite Equation~\eqref{eq:pat_rel_1d} as
 \begin{equation}
 \label{eq:t_func_1d}
 c_y(t) = f(x_i^1 + c_x(t)).
 \end{equation}
 We can now solve for $t$ in Equation~\eqref{eq:t_func_1d} in terms of $x_i^1$. If we substitute the solution $t(x_i)$ so obtained into Equation~\ref{eq:pat_rel_1d} we obtain 
\begin{equation}
\label{eq:h_func_1d}
 x_i^2 = h(x_i^1) = x_i^1 + c_x\circ t(x_i^1).
\end{equation}

\noindent
The area of the region $R$ depicted in Figure~\ref{fig:caval2} can, therefore, be computed as follows:
\begin{equation}
\int \limits_{\mathcal{C}} f(x)~dx = \int_a^b f \circ h (x)\, dx = \dfrac{1}{2}\int_1^4x\, dx = 3.75.  
\end{equation}
The expression for $h$ was found by employing Equation~\eqref{eq:pat_rel_1d}, Equation~\eqref{eq:t_func_1d} and Equation~\eqref{eq:h_func_1d}. Let us now verify that 
this result is correct by using a more conventional approach:
\begin{equation}
\int_0^2x\, dx+\int_2^44-x\, dx- \int_0^{\frac{1}{2}}x\, dx-\int_{\frac{1}{2}}^11-x\, dx = 3.75. 
\end{equation}

\noindent
In \cite{ackermann12}, the Cavalieri integral is denoted via $\int_{a(y)}^{b(y)} f(x)~dx$, where $a(y)$ denotes the functional representation of $\mathbf{c}(t) + \left <a,0 \right >$ and $b(y) = a(y)+(b-a)$

@@

\section{Double Cavalieri Integral}
In this section we define the double Cavalieri integral and we show that it can be used to calculate the volume of solids that can be decomposed into an infinite 
number of irregularly shaped regtangular infinitesimals.\\

@@

\noindent
Moreover, assume that $\mathbf{c}(t)$ is a translational curve with respect to $f(x,y)$ on $R$. Let us first partition $R$ into $n\times m$ subrectangles $R_{ij} = [x_{i},x_{i+1}] \times [y_{j},y_{j+1}]$ of equal width $\Delta x^1 = \Delta x_i^1 = x_{i+1}-x_i\ll (b-a)$ and length $\Delta y^1= \Delta y_j^1 = y_{j+1}-y_j\ll(d-c)$.
Furthermore, let the transformation $T$ map the $xy$--coordinate $(x_i^1,y_j^1)\in R$ (the lower left vertice of $R_{ij}$) at which $\mathbf{c}(t) + \left <x_i^1,y_j^1,0 \right >$ intersects the 
$xy$-plane to the $xy$--coordinate at which $\mathbf{c}(t) + \left <x_i^1,y_j^1,0 \right >$ intersects $f(x,y)$. Note, the domain of $T$ is $R$. Moreover, let us denote the range 
of $T$ with $S$. As the partition of $R$ can be made arbitrarily fine, $T:R\rightarrow S$ actually maps every point $(x_i^1,y_j^1)\in R$ to a point $(x_i^2,y_j^2)\in S$. In general, the transformation $T$ is defined as follows:
\begin{equation}
x_i^2 = h(x_i^1,y_j^1)~\textrm{and}~y_j^2 = s(x_i^1,y_j^1).
\end{equation}

@@

\noindent
Also, let $S_f = \{(x,y,z)\in\mathbb{R}^3|(x,y)\in S ~ \textrm{and} ~ z=f(x,y)\}$.  Let us further assume that $f(x,y) > 0$ for all $(x,y)\in S$. Let $\mathcal{S}$ denote the solid which is obtained by connecting the vertices of $R$ and 
$S_f$ via translations of $\mathbf{c}(t)$ along the $xy$ plane and then closing the resulting open faces that form due to the aforementioned connections by employing appropriate surfaces. Note that 
$\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals. As an example, the solid $\mathcal{S}$ depicted in Figure~\ref{fig:prismatoid_solid} is obtained if we set $f(x,y) = -x - y + 8$, $\mathbf{c}(t) = -\left<\frac{1}{2}t,t,t\right>$ and $R = \{(x,y)\in\mathbb{R}^2|0 \geq x \geq 4, 0 \geq x \geq 4\} \nonumber$.
Moreover, the domain $R$ and the range $S$ of the transformation function $T$ which is associated with the solid $\mathcal{S}$ is depicted in Figure~\ref{fig:prismatoid_regions}.  
In general, let $V_{\mathcal{S}}$ denote the volume of $\mathcal{S}$. We can approximate $V_{\mathcal{S}}$ with:
\begin{equation}
\label{eq:cav_sum_2d}
V_{\mathcal{S}} \approx \sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta x_i^1 \Delta y_j^1.
\end{equation}
Let us verify that this is indeed the case. If we closely inspect Equation~\ref{eq:cav_sum_2d} and Figure~\ref{fig:prismatoid_solid} we notice the following:
\begin{itemize}
\item $\Delta x_i^1$ and $\Delta y_j^1$ represent the width and the length of subrectangle $R_{ij}$.
\item The product $\Delta A = \Delta x_i^1\times \Delta y_j^1$ represents the area of subrectangle $R_{ij}$.
\item $f(x_i^2,y_j^2)$ represents the height of an irregular rectangular column or prism $\mathcal{K}_{ij}$ (or in the case of Figure~\ref{fig:prismatoid_solid} and oblique rectangular prism) whose base is $R_{ij}$. Furthermore, the lateral edges of this column are all translations 
of $\mathbf{c}(t)$ along the $xy$-plane. The cross-section between any parallel plane to the $xy$-plane and $\mathcal{K}_{ij}$ is a rectangle with the same dimensions as $R_{ij}$. 
\item The product $f(x_i^2,y_j^2)\times \Delta A$ represents the volume of $\mathcal{K}_{ij}$ (due to Cavalieri's principle).
\item The double Cavalieri sum $\sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta A$, therefore, approximates $V_{\mathcal{S}}$. 
\item In the limit the double Cavalieri sum approaches $V_{\mathcal{S}}$ (if this limit exists). 
\end{itemize}

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{prismatoid_solid.pdf}
\caption{The solid $\mathcal{S}$ which is bounded by the planes $f(x,y)=-x-y+8$, $2x$, $2(x-4)$, $y$, $y-4$ and the $xy$-plane. The figure also depicts the partition 
points $(x_i^1,y_i^1)\in\mathcal{R}$ and $(x_i^1,y_i^1)\in\mathcal{R}$ which is used in the double Cavalieri sum (see Equation~\eqref{eq:cav_2d}).}
\label{fig:prismatoid_solid}
\end{figure}

@@

\begin{eqnarray}
\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \lim_{n,m\rightarrow \infty} \sum_{j=1}^n\sum_{i=1}^n f(x_i^2,y_j^2) \Delta x_i^1\Delta y_j^1\\
&=&  \lim_{n,m\rightarrow \infty} \sum_{j=1}^n\sum_{i=1}^n f(h(x_i^1,y_j^1),s(x_i^1,y_j^1)) \Delta x_i^1\Delta y_j^1\\
&=& \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy.
\end{eqnarray}

\noindent
The only thing that now remains is to find expressions for $h$ and $s$. Note that, for every coordinate pair $(x_i^1,y_j^1)\in R$ and $(x_i^2,y_j^2)\in S$ there exists a unique $t\in \mathbb{R}$ such that 
\begin{equation}
\label{eq:vector_2}
\left< x_i^2, y_j^2, f(x_i^2,y_j^2) \right > = \left < x_i^1, y_j^1, 0 \right > + \left <c_x(t),c_y(t),c_z(t)\right >  
\end{equation}
The above equation can be re-expressed as:
\begin{equation}
\label{eq:c_z}
c_z(t) = f(x_i^1 + c_x(t),y_j^1 + c_y(t)). 
\end{equation}
We can now solve for $t$ in terms of $x_i^1$ and $y_j^1$. If we substitute this solution into Equation~\ref{eq:vector_2} we obtain:
\begin{equation}
\label{eq:h_2d}
x_i^2 = h(x_i^1,y_j^1) = x_i^1 + c_x\circ t(x_i^1,y_j^1) 
\end{equation}
and 
\begin{equation}
\label{eq:s_2d}
y_j^2 = s(x_i^1,y_j^1) = y_j^1 + c_y\circ t(x_i^1,y_j^1). 
\end{equation}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{prismatoid_real.jpg}
\caption{A physical not-to-scale model of the solid $\mathcal{S}$ depicted in Figure~\ref{fig:prismatoid_solid}.}
\label{fig:prismatoid_real}
\end{figure}

@@

\noindent
As we alluded to in the previous section we can calculate $V_{\mathcal{S}}$ using the double Cavalieri integral:
\begin{equation}
\label{eq:cav_ex1}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We can, however, only evaluate the above integral if the transformation $T$ is known (see Figure~\ref{fig:prismatoid_regions}). We can find $T$ using Equation~\eqref{eq:h_2d} and 
Equation~\eqref{eq:s_2d}. To evaluate Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} we first need to express $t$ in terms of $x_i^1$ and $y_j^1$. If we substitute $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t(x_i^1,y_j^1) = \frac{1}{5}x_i^1 + \frac{1}{5}y_j^1 + \frac{8}{5}. 
\end{equation}
Substituting $t(x_i^1,y_j^1)$ into Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} results in:
\begin{equation}
x_{i}^2 = h(x_{i}^1,y_{j}^1) = \frac{4}{5} x_{i}^1 - \frac{1}{5} y_{j}^1 + \frac{8}{5}  
\end{equation}
and
\begin{equation}
y_{j}^2 = r(x_{i}^1,y_{j}^1) = \frac{3}{5} x_{i}^1 - \frac{2}{5} y_{j}^1 + \frac{16}{5}.  
\end{equation}
We can now evaluate Equation~\eqref{eq:cav_ex1}:
\begin{eqnarray}
\label{eq:final_int_prismatoid}
V_{\mathcal{S}}=\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^4\int_0^4 f(h(x,y),r(x,y))~dydx\\
&=& \int_0^4 \int_0^4 -\frac{7}{5}x + \frac{3}{5} y + \frac{16}{5}~dydx\\
&=& 25.6
\end{eqnarray}

@@

\noindent
Since $\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals we can compute its volume using:
\begin{equation}
\label{eq:cav_ex2}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We, however, first need to calculate $T$ if we wish to evaluate Equation~\ref{eq:cav_ex2} (see Figure~\ref{fig:sphere_regions}). We can determine $T$ via 
Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d}. Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} in turn requires $t(x_i^1,y_j^2)$, which can be obtained via Equation~\eqref{eq:c_z}.
If we substite $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t = \frac{1}{4}[-2 x_{i}^1 - 2 y_{j}^1  -  1 + K(x_{i}^1,y_{j}^1)],
\end{equation}
where
\begin{equation}
K(x_{i}^1,y_{j}^1) = \sqrt{-4(x_{i}^1)^2+4x_{i}^1+8x_{i}^1y_{j}^1-4(y_{j}^1)^2+4y_{j}^1 + 33}. 
\end{equation}
As mentioned earlier, we can use $t(x_i^1,y_j^1)$ to find expressions for $h$ and $s$. The expressions of $h$ and $s$ so obtained can then be used to evaluate Equation~\eqref{eq:cav_ex2}:  
\begin{eqnarray}
\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^1\int_0^1 f(h(x,y),r(x,y))~dydx\\
&=& \int_0^1 \int_0^1 \sqrt{-(x+t(x,y))^2 - (y+t(x,y))^2 + 4} ~dydx\\
&=& 0.864043
\end{eqnarray}

>>>>

\section{Cavalieri Integral}
The single Cavalieri integral was originally defined by Ackermann et al. without having to resort to vector calculus \cite{ackermann12}. In this section, however, we will define the Cavalieri integral using vector calculus notation \cite{stewart1999}. As the reader will undoubtedly realize, if we define the single Cavalieri integral using vector calculus then it becomes trivial to
extend its definition into $\mathbb{R}^3$.\\

@@

\begin{definition}[Translational curve in $\mathbb{R}^2$]
If a curve $\mathbf{c}(t)$ in $\mathbb{R}^2$ intersects the $x$ axis at the origin and $\mathbf{c}(t) + \left < e,0 \right >$ intersects 
the function $f(x)$ at a single point for all $e\in [a,b]$ then $\mathbf{c}(t)$ is a translational curve 
with respect to $f(x)$ on $[a,b]$ \cite{ackermann12}. 
\end{definition}

@@

\begin{equation}
\label{eq:cav_1d}
\int \limits_{\mathcal{C}} f(x)~dx = \lim_{n \rightarrow \infty}  \sum_{i=0}^{n-1} f(x_i^2) \Delta x_i^1,
\end{equation}
with $\mathcal{C} = \{\mathbf{c}(t),[a,b]\}$. Equation~\eqref{eq:cav_1d} is difficult to evaluate in its current form. Luckily, the Cavalieri integral 
can be converted into an equivalent Riemann integral using the transformation $h$:
 \begin{equation}
 \int \limits_{\mathcal{C}} f(x)\, dx = \int_{a}^{b} f\circ h(x)~dx. \nonumber
\end{equation}

\noindent
The only thing that now remains is to find an expression for $h$. Note that, for every coordinate $(x_i^1,x_i^2) \in [a,b] \times [a',b']$ there exists a unique $t\in\mathbb{R}$ such that
\begin{equation}
 \label{eq:pat_rel_1d}
 \left <x_i^2,f(x_i^2) \right > = \left <x_i^1,0 \right> + \left<c_x(t),c_y(t) \right >.
 \end{equation}
 We can rewrite Equation~\eqref{eq:pat_rel_1d} as
 \begin{equation}
 \label{eq:t_func_1d}
 c_y(t) = f(x_i^1 + c_x(t)).
 \end{equation}
 We can now solve for $t$ in Equation~\eqref{eq:t_func_1d} in terms of $x_i^1$. If we substitute the solution $t(x_i)$ so obtained into Equation~\eqref{eq:pat_rel_1d} we obtain 
\begin{equation}
\label{eq:h_func_1d}
 x_i^2 = h(x_i^1) = x_i^1 + c_x\circ t(x_i^1).
\end{equation}

\noindent
Let us now compute the area of the region $R$ depicted in Figure~\ref{fig:caval2} using the Cavalieri integral:
\begin{equation}
\int \limits_{\mathcal{C}} f(x)~dx = \int_a^b f \circ h (x)\, dx = \dfrac{1}{2}\int_1^4x\, dx = 3.75. \nonumber 
\end{equation}
The expression for $h$ was found by employing Equation~\eqref{eq:pat_rel_1d}, Equation~\eqref{eq:t_func_1d} and Equation~\eqref{eq:h_func_1d}. Let us now verify that 
this result is correct by using a more conventional approach:
\begin{equation}
\int_0^2x\, dx+\int_2^44-x\, dx- \int_0^{\frac{1}{2}}x\, dx-\int_{\frac{1}{2}}^11-x\, dx = 3.75. \nonumber
\end{equation}

\noindent
In \cite{ackermann12}, the Cavalieri integral is denoted via $\int_{a(y)}^{b(y)} f(x)~dx$, where $a(y)$ denotes the functional representation of $\mathbf{c}(t) + \left <a,0 \right >$ and $b(y) = a(y)+(b-a)$.

@@

\section{Double Cavalieri Integral}
In this section we extend the definition of the Cavalieri integral into $\mathbb{R}^3$, i.e. we define the double Cavalieri integral. We also show that it can be used to calculate the volume of solids that can be decomposed into an infinite 
number of irregularly shaped regtangular infinitesimals.\\

@@

\noindent
Moreover, assume that $\mathbf{c}(t)$ is a translational curve with respect to $f(x,y)$ on $R$. Let us first partition $R$ into $n\times m$ subrectangles $R_{ij} = [x_{i},x_{i+1}] \times [y_{j},y_{j+1}]$ of equal width $\Delta x^1 = \Delta x_i^1 = x_{i+1}-x_i\ll (b-a)$ and length $\Delta y^1= \Delta y_j^1 = y_{j+1}-y_j\ll(d-c)$.
Furthermore, let the transformation $T$ map the $xy$--coordinate $(x_i^1,y_j^1)\in R$ (the lower left vertice of $R_{ij}$) at which $\mathbf{c}(t) + \left <x_i^1,y_j^1,0 \right >$ intersects the 
$xy$-plane to the $xy$--coordinate at which $\mathbf{c}(t) + \left <x_i^1,y_j^1,0 \right >$ intersects $f(x,y)$. Note, the domain of $T$ is $R$. Moreover, let us denote the range 
of $T$ with $S$. As the partition of $R$ can be made arbitrarily fine, $T:R\rightarrow S$ actually maps every point $(x_i^1,y_j^1)\in R$ to a point $(x_i^2,y_j^2)\in S$. In general, the transformation $T$ is defined as follows:
\begin{equation}
x_i^2 = h(x_i^1,y_j^1)~\textrm{and}~y_j^2 = s(x_i^1,y_j^1).\nonumber
\end{equation}

@@

\noindent
Also, let $S_f = \{(x,y,z)\in\mathbb{R}^3|(x,y)\in S ~ \textrm{and} ~ z=f(x,y)\}$.  Let us further assume that $f(x,y) > 0$ for all $(x,y)\in S$. Let $\mathcal{S}$ denote the solid which is obtained by connecting the vertices of $R$ and 
$S_f$ via translations of $\mathbf{c}(t)$ along the $xy$ plane and then closing the resulting open faces that form due to the aforementioned connections by employing appropriate surfaces. Note that 
$\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals. As an example, the solid $\mathcal{S}$ depicted in Figure~\ref{fig:prismatoid_solid} is obtained if we set $f(x,y) = -x - y + 8$, $\mathbf{c}(t) = -\left<\frac{1}{2}t,t,t\right>$ and $R = \{(x,y)\in\mathbb{R}^2|0 \geq x \geq 4, 0 \geq x \geq 4\} \nonumber$.
Moreover, the domain $R$ and the range $S$ of the transformation function $T$ which is associated with the solid $\mathcal{S}$ is depicted in Figure~\ref{fig:prismatoid_regions}.  
In general, let $V_{\mathcal{S}}$ denote the volume of $\mathcal{S}$. We can approximate $V_{\mathcal{S}}$ with:
\begin{equation}
\label{eq:cav_sum_2d}
V_{\mathcal{S}} \approx \sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta x_i^1 \Delta y_j^1.
\end{equation}
Let us verify that this is indeed the case. If we closely inspect Equation~\eqref{eq:cav_sum_2d} and Figure~\ref{fig:prismatoid_solid} we notice the following:
\begin{itemize}
\item $\Delta x_i^1$ and $\Delta y_j^1$ represent the width and the length of subrectangle $R_{ij}$.
\item The product $\Delta A = \Delta x_i^1\times \Delta y_j^1$ represents the area of subrectangle $R_{ij}$.
\item $f(x_i^2,y_j^2)$ represents the height of an irregular rectangular column or prism $\mathcal{K}_{ij}$ (or in the case of Figure~\ref{fig:prismatoid_solid} and oblique rectangular prism) whose base is $R_{ij}$. Furthermore, the lateral edges of this column are all translations 
of $\mathbf{c}(t)$ along the $xy$-plane. The cross-section between any parallel plane to the $xy$-plane and $\mathcal{K}_{ij}$ is a rectangle with the same dimensions as $R_{ij}$. 
\item The product $f(x_i^2,y_j^2)\times \Delta A$ represents the volume of $\mathcal{K}_{ij}$ (due to Cavalieri's principle).
\item The double Cavalieri sum $\sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta A$, therefore, approximates $V_{\mathcal{S}}$. 
\item In the limit the double Cavalieri sum approaches $V_{\mathcal{S}}$ (if this limit exists). 
\end{itemize}

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{prismatoid_solid.pdf}
\caption{The solid $\mathcal{S}$ which is bounded by the planes $f(x,y)=-x-y+8$, $2x$, $2(x-4)$, $y$, $y-4$ and the $xy$-plane. The figure also depicts the partition 
points $(x_i^1,y_j^1)\in\mathcal{R}$ and $(x_i^2,y_j^2)\in\mathcal{S}$ which is used in the double Cavalieri sum (see Equation~\eqref{eq:cav_sum_2d}).}
\label{fig:prismatoid_solid}
\end{figure}

@@

\begin{eqnarray}
\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \lim_{n,m\rightarrow \infty} \sum_{j=1}^n\sum_{i=1}^n f(x_i^2,y_j^2) \Delta x_i^1\Delta y_j^1\nonumber\\
&=&  \lim_{n,m\rightarrow \infty} \sum_{j=1}^n\sum_{i=1}^n f(h(x_i^1,y_j^1),s(x_i^1,y_j^1)) \Delta x_i^1\Delta y_j^1\nonumber\\
&=& \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy.\nonumber
\end{eqnarray}

\noindent
The only thing that now remains is to find expressions for $h$ and $s$. Note that, for every coordinate pair $(x_i^1,y_j^1)\in R$ and $(x_i^2,y_j^2)\in S$ there exists a unique $t\in \mathbb{R}$ such that 
\begin{equation}
\label{eq:vector_2}
\left< x_i^2, y_j^2, f(x_i^2,y_j^2) \right > = \left < x_i^1, y_j^1, 0 \right > + \left <c_x(t),c_y(t),c_z(t)\right >  
\end{equation}
The above equation can be re-expressed as:
\begin{equation}
\label{eq:c_z}
c_z(t) = f(x_i^1 + c_x(t),y_j^1 + c_y(t)). 
\end{equation}
We can now solve for $t$ in terms of $x_i^1$ and $y_j^1$. If we substitute this solution into Equation~\eqref{eq:vector_2} we obtain:
\begin{equation}
\label{eq:h_2d}
x_i^2 = h(x_i^1,y_j^1) = x_i^1 + c_x\circ t(x_i^1,y_j^1) 
\end{equation}
and 
\begin{equation}
\label{eq:s_2d}
y_j^2 = s(x_i^1,y_j^1) = y_j^1 + c_y\circ t(x_i^1,y_j^1). 
\end{equation}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.4\textwidth]{prismatoid_real.jpg}
\caption{A physical not-to-scale model of the solid $\mathcal{S}$ depicted in Figure~\ref{fig:prismatoid_solid}.}
\label{fig:prismatoid_real}
\end{figure}

@@

\noindent
As we alluded to in the previous section we can calculate $V_{\mathcal{S}}$ using the double Cavalieri integral:
\begin{equation}
\label{eq:cav_ex1}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We can, however, only evaluate the above integral if the transformation $T$ is known (see Figure~\ref{fig:prismatoid_regions}). We can find $T$ using Equation~\eqref{eq:h_2d} and 
Equation~\eqref{eq:s_2d}. To evaluate Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} we first need to express $t$ in terms of $x_i^1$ and $y_j^1$. If we substitute $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t(x_i^1,y_j^1) = \frac{1}{5}x_i^1 + \frac{1}{5}y_j^1 + \frac{8}{5}. \nonumber
\end{equation}
Substituting $t(x_i^1,y_j^1)$ into Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} results in:
\begin{equation}
x_{i}^2 = h(x_{i}^1,y_{j}^1) = \frac{4}{5} x_{i}^1 - \frac{1}{5} y_{j}^1 + \frac{8}{5}  \nonumber
\end{equation}
and
\begin{equation}
y_{j}^2 = s(x_{i}^1,y_{j}^1) = \frac{3}{5} x_{i}^1 - \frac{2}{5} y_{j}^1 + \frac{16}{5}.  \nonumber
\end{equation}
We can now evaluate Equation~\eqref{eq:cav_ex1}:
\begin{eqnarray}
\label{eq:final_int_prismatoid}
V_{\mathcal{S}}=\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^4\int_0^4 f(h(x,y),s(x,y))~dydx\nonumber\\
&=& \int_0^4 \int_0^4 -\frac{7}{5}x + \frac{3}{5} y + \frac{16}{5}~dydx\nonumber\\
&=& 25.6.\nonumber
\end{eqnarray}

@@

\noindent
Since $\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals we can compute its volume using:
\begin{equation}
\label{eq:cav_ex2}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We, however, first need to calculate $T$ if we wish to evaluate Equation~\eqref{eq:cav_ex2} (see Figure~\ref{fig:sphere_regions}). We can determine $T$ via 
Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d}. Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} in turn requires $t(x_i^1,y_j^2)$, which can be obtained via Equation~\eqref{eq:c_z}.
If we substite $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t = \frac{1}{4}[-2 x_{i}^1 - 2 y_{j}^1  -  1 + K(x_{i}^1,y_{j}^1)],\nonumber
\end{equation}
where
\begin{equation}
K(x_{i}^1,y_{j}^1) = \sqrt{-4(x_{i}^1)^2+4x_{i}^1+8x_{i}^1y_{j}^1-4(y_{j}^1)^2+4y_{j}^1 + 33}.\nonumber
\end{equation}
As mentioned earlier, we can use $t(x_i^1,y_j^1)$ to find expressions for $h$ and $s$. The expressions of $h$ and $s$ so obtained can then be used to evaluate Equation~\eqref{eq:cav_ex2}:  
\begin{eqnarray}
\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^1\int_0^1 f(h(x,y),s(x,y))~dydx\nonumber\\
&=& \int_0^1 \int_0^1 \sqrt{-(x+t(x,y))^2 - (y+t(x,y))^2 + 4} ~dydx\nonumber\\
&=& 0.864043.\nonumber
\end{eqnarray}
